Dynamics of Laterally Propagating Flames in X-ray Bursts

### 2019-10-14 xrb/paper.tex
write section on the sponge

\AssignTo{Mike, Alice}

@@

For deflagrations, we either need to resolve the structure of the
reaction zone or use a flame model.  Flame models usually assume that
the flame structure is thin compared to the size of the system (see,
e.g., \citet{Ropke2007} for applications to Type Ia supernovae).  For
XRBs however, the flame thickness is comparable to the scale height of
the atmosphere, so we cannot use these approximations.  Accurate
models of flames therefore require that we resolve the thermal width,
which is $\mathcal{O}(10~cm)$ for helium flames~\citep{Timmes00}.

@@

 An equation of state of the form $p = p(\rho, e,
\{X_k\})$ completes the thermodynamic description of the
system.  Finally, $\kth$ is the thermal conductivity and $T$ is the
temperature.  Diffusion is treated explicitly in time.

Since the neutron star rotates, we work in a corotating frame, with
angular velocity $\Omegab$.  Further, for the two-dimensional
simulations presented here, we work in axisymmetric coordinates,
but we advect a third component of velocity, coming out of the
simulation plane, that participates in the Coriolis force (sometimes
described as a 2.5D simulation).  We will take the
\castro\ $x$-coordinate to be the cylindrical radial coordinate with
corresponding velocity $u$, the \castro\ $y$-coordinate to be the
cylindrical vertical coordinate with corresponding velocity $v$, and the
\castro\ $z$-coordinate to be the cylindrical azimuthal coordinate,
with corresponding velocity $w$.  A righthanded coordinate system has
positive $w$ pointing out of the page.  We will take $\Omegab =
\Omega_0 \hat{\bf y}$ for the angular rotation rate, and $\gb = -g
\hat{\bf y}$ for the gravitational acceleration, with $g$ constant.
With these choices, the Coriolis force is:
\begin{equation}
-2\rho \Omegab \times \Ub =
   -2\rho \left ( \Omega_0 w \hat{x} - \Omega_0 u \hat{z} \right )
\end{equation}
We will neglect the centrifugal force---with our plane-parallel
geometry, this will act only in the lateral direction, and is not
expected to greatly affect the dynamics.  Carrying the Coriolis force
allows us to capture the geostrophic balance\MarginPar{ref, Spitkovsky 2002?} that sets
up via lateral hydrostatic equilibrium.

@@

sponge\AssignTo{Mike will write}

In the discussions below, we will use the Castro coordinate
names, $(x, y)$ in our notation.

>>>>

@@

For deflagrations, we either need to resolve the structure of the
reaction zone or use a flame model.  Flame models usually assume that
the flame structure is thin compared to the size of the system (see,
e.g., \citet{Ropke2007} for applications to Type Ia supernovae).  For
XRBs however, the flame thickness is comparable to the scale height of
the atmosphere, so we cannot use these approximations.  Accurate
models of flames therefore require that we resolve the thermal width,
which is $\mathcal{O}(10~\mbox{cm})$ for helium flames~\citep{Timmes00}.

@@

An equation of state of the form $p = p(\rho, e, X_k)$ completes the
thermodynamic description of the system.  Finally, $\kth$ is the
thermal conductivity and $T$ is the temperature.  Diffusion is treated
explicitly in time.

Since the neutron star rotates, we work in a corotating frame, with
angular velocity $\Omegab$.  Further, for the two-dimensional
simulations presented here, we work in axisymmetric coordinates,
but we advect a third component of velocity, coming out of the
simulation plane, that participates in the Coriolis force (sometimes
described as a 2.5D simulation).  We will take the
\castro\ $x$-coordinate to be the cylindrical radial coordinate with
corresponding velocity $u$, the \castro\ $y$-coordinate to be the
cylindrical vertical coordinate with corresponding velocity $v$, and the
\castro\ $z$-coordinate to be the cylindrical azimuthal coordinate,
with corresponding velocity $w$.  A righthanded coordinate system has
positive $w$ pointing out of the page.  We will take $\Omegab =
\Omega_0 \hat{\bf y}$ for the angular rotation rate, and $\gb = -g
\hat{\bf y}$ for the gravitational acceleration, with $g$ constant.
With these choices, the Coriolis force is:
\begin{equation}
-2\rho \Omegab \times \Ub =
   -2\rho \left ( \Omega_0 w \hat{x} - \Omega_0 u \hat{z} \right )
\end{equation}
We will neglect the centrifugal force---with our plane-parallel
geometry, this will act only in the lateral direction, and is not
expected to greatly affect the dynamics.  Carrying the Coriolis force
allows us to capture the geostrophic balance\MarginPar{ref, Spitkovsky 2002?} that sets
up via lateral hydrostatic equilibrium.  In the discussions below, we will use the Castro coordinate
names, $(x, y)$ in our notation.

@@

When we begin the simulation, there is a transient phase as the flame
gets established.  Material that is forced upward will encounter the
steep density gradient at the top of the atmosphere and accelerate as
it is blown out of the atmosphere.  Eventually this material will fall
back to the top of the atmosphere.  In this paper, we are mostly
concerned with the behavior of the flame and not any material that is
violently blown out of the top of the atmosphere, so we apply a sponge
to this region.  This is similar to what we've done in \cite{xrb3d}.
The sponge drives the velocity of the material in the low
density \MarginPar{add mathematical form of sponge} regions at the top
of the atmosphere to zero.  This sponging helps increase our timestep
as well.
